---
title: "Covid-19 Report"
output:
  pdf_document: default
  html_document: default
---


## 1 Introduciton

this data set ...

Libraries needed:

```{r library, message= FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

Importing the data:

```{r ImportData, message= FALSE}

url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"
file_names <- c("time_series_covid19_confirmed_global.csv","time_series_covid19_deaths_global.csv", "time_series_covid19_confirmed_US.csv", "time_series_covid19_deaths_US.csv"  )
urls <- str_c(url_in, file_names)
uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"

global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])
US_cases <- read_csv(urls[3])
US_deaths <- read_csv(urls[4])
uid = read_csv(uid_lookup_url) %>%
  select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))
```

## 2 Tidying the Data

Here we are cleaning up the dataset by changing appropriate types and getting rid of any columns not needed.

```{r tidy_global data}
global_cases <- global_cases %>% 
  pivot_longer(cols = -c('Province/State',
                         'Country/Region', Lat, Long),
               names_to = "date", values_to = "cases") %>%
  select(-c(Lat, Long))

global_deaths <- global_deaths %>% 
  pivot_longer(cols = -c('Province/State',
                         'Country/Region', Lat, Long),
               names_to = "date", values_to = "deaths") %>%
  select(-c(Lat, Long))

global <- global_cases %>%
  full_join(global_deaths) %>%
  rename(Country_Region = 'Country/Region',
         Province_State = 'Province/State') %>%
  mutate(date = mdy(date))

global <- global %>% filter(cases>0)

global <- global %>%
  unite("Combined_Key",
         c(Province_State, Country_Region),
         sep = ", ",
         na.rm = TRUE,
         remove = FALSE)

global <- global %>%
  left_join(uid, by = c("Province_State", "Country_Region")) %>%
  select(-c(UID, FIPS)) %>%
  select(Province_State, Country_Region, date, cases, deaths, Population, Combined_Key)

global

```

```{r US data}

US_cases <- US_cases %>%
  pivot_longer(cols = -(UID:Combined_Key),
               names_to = "date",
               values_to = "cases") %>%
  select(Admin2:cases) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat, Long_))

US_deaths <- US_deaths %>%
  pivot_longer(cols = -(UID:Population),
               names_to = "date",
               values_to = "deaths") %>%
  select(Admin2:deaths) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat, Long_))

US <- US_cases %>%
  full_join(US_deaths)

US

```

##  3 Visualization and Analysis

Let start by lookinga at ...

```{r Visualizing_US}
US_by_State <- US %>%
  group_by(Province_State, Country_Region,date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths),
            Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths *1000000 / Population) %>%
  select(Province_State, Country_Region, date, cases, deaths, deaths_per_mill, Population) %>%
  ungroup()

US_total <- US_by_State %>%
  group_by(Country_Region,date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths),
            Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths *1000000 / Population) %>%
  select(Country_Region, date, cases, deaths, deaths_per_mill, Population) %>%
  ungroup()

US_total %>%
  filter(cases > 0) %>%
  ggplot(aes(x = date, y = cases))+
  geom_line(aes(color = "cases"))+
  geom_point(aes(color = "cases"))+
  geom_line(aes(y = deaths, color = "deaths"))+
  geom_point(aes(y = deaths, color = "deaths"))+
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text = element_text(angle = 90))+
  labs(title = "Covid-19 in US", y = NULL)

```
```{r state Visual}

state <- "New York" 
US_by_State%>%
  filter(Province_State == state)%>%
  filter(cases > 0) %>%
  ggplot(aes(x = date, y = cases))+
  geom_line(aes(color = "cases"))+
  geom_point(aes(color = "cases"))+
  geom_line(aes(y = deaths, color = "deaths"))+
  geom_point(aes(y = deaths, color = "deaths"))+
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text = element_text(angle = 90))+
  labs(title = "Covid-19 in New York", y = NULL)

```
```{r New_Cases}

US_by_State <- US_by_State %>%
  mutate(new_cases = cases - lag(cases),
         new_deaths = deaths - lag(deaths))
US_total <- US_total %>%
  mutate(new_cases = cases - lag(cases),
         new_deaths = deaths - lag(deaths))  

```

``` {r NewCasesGraphUS}
US_total %>%
  filter(cases > 0) %>%
  ggplot(aes(x = date, y = new_cases))+
  geom_line(aes(color = "new_cases"))+
  geom_point(aes(color = "new_cases"))+
  geom_line(aes(y = new_deaths, color = "new_deaths"))+
  geom_point(aes(y = new_deaths, color = "new_deaths"))+
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text = element_text(angle = 90))+
  labs(title = "New Cases Covid-19 in US", y = NULL)

``` 

``` {r NewCasesGraphState}

state <- "New York" 
US_by_State%>%
  filter(Province_State == state)%>%
  filter(cases > 0) %>%
  ggplot(aes(x = date, y = new_cases))+
  geom_line(aes(color = "new_cases"))+
  geom_point(aes(color = "new_cases"))+
  geom_line(aes(y = new_deaths, color = "new_deaths"))+
  geom_point(aes(y = new_deaths, color = "new_deaths"))+
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text = element_text(angle = 90))+
  labs(title = "Covid-19 in New York", y = NULL)

```

```{r State_Comparision}

US_state_total <- US_by_State %>%
  group_by(Province_State) %>%
  summarise(deaths = max(deaths), cases = max(cases), population = max(Population),
            cases_per_thou = 1000* cases / population,
            deaths_per_thou = 1000* deaths / population) %>%
  filter(cases> 0, population > 0)
```

```{r Best_States}
US_state_total %>%
  slice_min(deaths_per_thou, n=10) %>%
select(deaths_per_thou, cases_per_thou, everything())

```

```{r Worse_States}
US_state_total %>%
  slice_max(deaths_per_thou, n=10) %>%
select(deaths_per_thou, cases_per_thou, everything())

```
## 4 Modeling

```{r Modeling}
mod <- lm(deaths_per_thou ~ cases_per_thou, data = US_state_total)
summary(mod)

US_state_total %>% slice_min(cases_per_thou)
US_state_total %>% slice_max(cases_per_thou)

#x_grid <- seq(1, 151)
#new_df <- tibble(cases_per_thou = x_grid)
#US_state_total %>% mutate(pred = predict(mod))
US_tot_w_pred <- US_state_total %>% mutate(pred = predict(mod))
US_tot_w_pred

US_tot_w_pred %>% ggplot()+
  geom_point(aes(x = cases_per_thou, y = deaths_per_thou), color = "blue")+
  geom_point(aes(x = cases_per_thou, y = pred), color = "red")

```

## 5 Conclusion

In conclusion, our analysis reveals a highly significant correlation between Total Shootings and Total Deaths, indicating a strong linear relationship with low variance. This suggests that Total Shootings is a strong predictor of Total Deaths. However, to enhance the predictive accuracy of the model, incorporating additional factors could be beneficial.

It is important to acknowledge potential limitations in the data. Difference in the data collection methods across different precincts could introduce bias into the resutls. Biased reporting or underreporting in certain ares/ demographics could influence the results. The removal of any rows that have NA may also introduce bias into the results.


depends on the variable used 
how the variables were measured
surveys - how the questions were phased
who you ask, who asked them 
how do you handle outliers
are the death reported all the same?



